%        File: thesis.tex
%     Created: Tue Jan 12 11:00 AM 2010 C
% Last Change: Tue Jan 12 11:00 AM 2010 C
%
\documentclass[review]{ipgthesis}
%\documentclass[a4paper]{lcavthesis}

% All external packages
\input{packages}

% Theorem-like definitions
\input{theoremdefs}

% All macros and definitions
\input{thesismac}

% Style definitions for TikZ
\input{tikzstyles}

% Various commands affecting the whole document
\bibliographystyle{IEEEtranS}
\frenchspacing

\begin{document}
\setcounter{chapter}{1}
\section{Towards a General SDR Upper Bound}\label{sec:gensdrub}

The preceding sections used a bounding technique due to Ziv
that made it possible to lower bound the MSE for a given encoder regardless of
the decoder. This technique works only if the encoding function fulfills certain
criteria. For example, the proofs of the lower bounds in Section~2.2 worked
because by a smart choice of~$\Delta$, the distance $\|f(s) - f(s+\Delta)\|$
could be upper bounded for all~$s$ in the interval $[A,B-\Delta)$.

More generally, suppose there exists a set $\Xi \subseteq \R$ and a
differentiable function~$\xi$ such that $\pS(s) \ge \pmin > 0$ for all $s \in
\Xi \cup \xi(\Xi)$.  Then the MSE can be bounded as
\begin{equation*}
  \mse \ge \pmin \int_\Xi \msecond ds
\end{equation*}
and
\begin{align*}
  \mse &\ge \pmin \int_{\xi(\Xi)} \msecond ds \\
  &= \pmin \int_\Xi \msecondxi \xi'(s) ds.
\end{align*}
Combining the two bounds results in
\begin{equation*}
  \mse \ge \frac{\pmin}{2} \int_{\Xi} \left( \msecond + \msecondxi \xi'(s)
  \right) ds.
\end{equation*}
For example, letting $\Xi = [A, B-\Delta)$ and $\xi(s) = s + \Delta$ implies
$\xi'(s) = 1$ and thus
\begin{equation*}
  \mse \ge \frac{\pmin}{2} \intabd \left( \msecond + \msecondd \right) ds,
\end{equation*}
which is exactly the bound that occurs in the proof for Ziv's bound. 

Unfortunately, it seems that if $\xi'(s) \ne 1$, then we cannot use the trick
with Markov's inequality to lower bound this with the binary error probability.
We would get something like
\begin{multline*}
  \mse \ge \frac c2 \int_\Xi (s - \xi(s))^2 \left(
  \Pr\left[|\Sh - S| \ge \frac{s - \xi(s)}{2} \mid s \right] \right. + \\
  \left. \Pr\left[|\Sh - S| \ge \frac{s - \xi(s)}{2} \mid \xi(s) \right] \xi'(s)
  \right) ds.
\end{multline*}
The factor $\xi'(s)$ makes it difficult to lower bound this with binary error
probability. 

On the other hand, $\xi'(s) = 1$ implies $\xi(s) = s + c$, and we are back to
the original case.

\end{document}
